### 2013-10-10 Response to referees.txt
Nearly ready to resubmit

We thank the referees for their detailed reports.  We have addressed their 
concerns and clarified answers to their questions in a new version of the 
manuscript.

Response to suggestions of referee 1:

1. We have clarified the wording of the discussion at the bottom of page 4, to
better explain the derivation of Eq. (12), which is the reduced, approximate 
form of (11).

2. We had indeed overlooked defining \tau. We have now added the definition
of \tau, and renamed the period of cyclic variations from \tau to T to
avoid confusion.

3. We have also added a paragraph at the end of Section 3 clearly explaining
our definition of number squeezing in the multimode case, and refer to this
definition in the caption of Figure 4.

4. We thank the referee for bring the extra reference on atomic Kerr squeezing
to our attention, and we have added it to our paper.

Response to questions and suggestions of referee 2:

1.  We have added a discussion of the practicality of using squeezing to improve a precision 
measurement relative to the possibility of increasing the atom number.  In fact, these two methods 
should ideally be used together if possible, as they are essentially entirely compatible and 
independent improvements to the sensitivity.  We include the highly relevant reference that the 
referee quoted.

2.  We have added a comment that the large number, box-potential geometries are most obviously
adapted to precision inertial measurements rather than in-trap measurements with goals of spatial 
resolution.

3.  We have rewritten our discussion of the effect of trap geometry on squeezing in response to 
the referee's very reasonable comments that it was confusing and apparently contradictory.  The 
effect of dimensionality changes character completely depending on whether the density is kept 
constant, or the total atom number is kept constant.  We have elaborated on this point, and made
the manuscript clear on this key issue.

     These two conclusions seem contradictory, and it is not clear how to relate them. This is
    made more difficult because figure 10 is missing the part labelled "Right", which as I
    understand should show the achievable squeezing in the 3D case. The authors appear to
    resolve this seeming contradiction in the conclusion, but it is unclear from the text how
    this relates to the parameters used in the different simulations.